---
title: "Mining mega-diverse genera of flowering plants"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining mega-diverse genera of flowering plants}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.table{
  width: auto;
  font-size: 14px;
}
.table caption {
  font-size: 1em;
}
```

Here in this article, we show how to use the package's function `megaGen` for 
mining the mega-diverse genera for any family of flowering plants. [Frodin (2004)](https://doi.org/10.2307/4135449) suggests 500 species as the threshold 
for a genus to be considered mega-diverse. The function `megaGen` 
uses any integer numeric value in the argument `thld` to define the threshold 
for a mega-diverse genus.\

\

# Setup

Install the latest development version of __expowo__ from 
[GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("DBOSlab/expowo")
```
```{r package load, message=FALSE, warning=FALSE}
library(expowo)
```

\

## Mining mega-diverse genera for any plant family

The function `megaGen` uses a specified threshold number to return a data frame 
or save a CSV file listing all mega-diverse genera for any family of flowering 
plants and their associated number of accepted species. In this example, we 
performed a search for three plant families using a threshold of 500 species, 
but the output only contains results for **Aristolochia** and **Begonia** of 
__Aristolochiaceae__ and __Begoniaceae__ families.\

```{r, eval = FALSE}
ABM <- megaGen(family = c("Aristolochiaceae", "Begoniaceae", "Martyniaceae"),
               thld = 500,
               verbose = TRUE,
               save = FALSE,
               dir = "results_megaGen",
               filename = "Arist_Begon_Martyniaceae_search")
```

\

```{r, echo = FALSE, warning = FALSE}
utils::data("angioGenera")
family <- c("Aristolochiaceae", "Begoniaceae")
genus <- c("Aristolochia", "Begonia")
species_number <- angioGenera$species_number[angioGenera$genus %in% genus]
authors <- c("L.", "L.")
scientific_name <- c("Aristolochia L.", "Begonia L.")
powo_uri <- angioGenera$powo_uri[angioGenera$genus %in% genus]
kew_id <- angioGenera$kew_id[angioGenera$genus %in% genus]

res <- data.frame(family, genus, authors, scientific_name,
                  species_number, kew_id, powo_uri)

knitr::kable(res,
             row.names = FALSE,
             caption = "TABLE 1. A general `megaGen` search to mining the
             mega-diverse genera for some specific angiosperm families,
             based on a specified threshold of 500 species.")
```

\

## Mining mega-diverse genera accross all angiosperm families

To mine a global checklist of mega-diverse genera for all families of flowering 
plants, including their associated species number, we recommend to load the 
dataframe-formatted data object called `POWOcodes` that comes associated with
the __expowo__ package. The `POWOcodes` data object already contains the 
URI addresses for all angiosperms families recognized in the 
[POWO](https://powo.science.kew.org) database, you just need to call it to the
R environment.\

The example below shows how to mine a global checklist of all most speciose 
genera of flowering plants by using the vector of all angiosperm families and 
associated URI addresses stored in the `POWOcodes` object.\

```{r, eval = FALSE}
data(POWOcodes)

ALL_megaGen <- megaGen(POWOcodes$family,
                       thld = 500,
                       verbose = TRUE,
                       save = FALSE,
                       dir = "results_megaGen",
                       filename = "all_megadiverse_angiosperm_genera")
```

\

## Reference

POWO (2019). "Plants of the World Online. Facilitated by the Royal Botanic Gardens, Kew. Published on the Internet; http://www.plantsoftheworldonline.org/ Retrieved April 2023."
