---
title: "Mining accepted genera for any angiosperm family"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining accepted genera for any angiosperm family}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The main goal of the __expowo__ package is to retrieve information about the 
diversity and distribution of any plant family as publicly available at the 
taxonomically verified POWO database [Plants of the World Online (POWO)](https://powo.science.kew.org).\

The package is intended to efficiently mine the content within the source html pages
for any specific genus and family. It can return a comma-separated values (CSV) 
file with the number of accepted species and country-level distribution for any genus
as well as the full checklist of accepted species in any genus or family, their authorship, original publication and global distribution.\

Here in this article, we show how to use the package's function `powoGenera` for 
mining all accepted genera for any family of flowering plants.

# Setup

__expowo__ is not on CRAN yet but you can install the latest development version from [GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("deborazuanny/expowo")
```
```{r package load}
library(expowo)
```

# Mining all accepted genera for any angiosperm family

The function `powoGenera` returns a dataframe or saves a CSV file listing all 
genera with associated number of accepted species and their global geographic 
distribution at country or botanical level. The global classification of botanical 
divisions follows the [World Geographical Scheme](https://www.tdwg.org/standards/wgsrpd/) for Recording Plant Distributions, 
which is already associated each taxon's distribution in the POWO.

```{r, eval = FALSE}
library(taxize)
fam <- c("Aristolochiaceae", "Begoniaceae", "Lecythidaceae", "Martyniaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

powoGenera(powocodes$family, powocodes$uri,
           verbose = FALSE,
           save = TRUE,
           dir = "results_powoGenera/",
           filename = "Arist_Begon_Lecyt_Martyniaceae_search")
```

```{r, echo = FALSE, warning = FALSE}
library(taxize)
fam <- c("Aristolochiaceae", "Begoniaceae", "Lecythidaceae", "Martyniaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam, messages = FALSE)))

res <- powoGenera(powocodes$family, powocodes$uri,
                  verbose = FALSE,
                  save = FALSE)

knitr::kable(res[-c(7:10)])
```

# Narrowing down the `powoGenera` search based on a specified country list

You can also narrow down the search for all accepted genera of any family with 
focus on just a particular country or a list of countries. You only need to define 
a vector of country names in the argument \code{country}. In the example below,
see that we have originally searched for the genera within the families __Aristolochiaceae__,
__Lecythidaceae__, __Fagaceae__, and __"Dipterocarpaceae__, but the function 
only returned a list of genera of __Aristolochiaceae__ and __Lecythidaceae__,
because the remaining families do not have any genera recorded in the list
of countries of interest, i.e. __Argentina__, __Brazil__ or __French Guiana__.

```{r, eval = FALSE}
library(taxize)
fam <- c("Aristolochiaceae", "Lecythidaceae", "Fagaceae", "Dipterocarpaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

powoGenera(powocodes$family, powocodes$uri,
           country = c("Argentina", "Brazil", "French Guiana"),
           verbose = FALSE,
           save = TRUE,
           dir = "results_powoGenera/",
           filename = "country_constrained_search")
```

```{r, echo = FALSE, warning = FALSE}
library(taxize)
fam <- c("Aristolochiaceae", "Lecythidaceae", "Fagaceae", "Dipterocarpaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam, messages = FALSE)))

res <- powoGenera(powocodes$family, powocodes$uri,
                  country = c("Argentina", "Brazil", "French Guiana"),
                  verbose = FALSE,
                  save = FALSE)

knitr::kable(res[-c(7:10)])
```


# Narrowing down the `powoGenera` search based on a specified genus list

You may want to retrieve information for just one or a list of accepted 
genera from a given country (or from a list of countries). Just like before,
you only need to define a vector of genus names in the argument \code{genus} and
a vector of country names in the argument \code{country}. In the example below,
see that we have again searched for just the genera ***Asarum*** and ***Bertholletia***
of the families __Aristolochiaceae__ and __Lecythidaceae__, but the function 
only returned the Lecythidaceae genus ***Bertholletia***, because ***Asarum*** 
does not occur in any of the provided list of countries, i.e. __Argentina__, __Brazil__ 
or __French Guiana__.

```{r, eval = FALSE}
library(taxize)
fam <- c("Aristolochiaceae", "Lecythidaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

powoGenera(powocodes$family, powocodes$uri,
           genus = c("Asarum", "Bertholletia"),
           country = c("Argentina", "Brazil", "French Guiana"),
           verbose = TRUE,
           save = TRUE,
           dir = "results_powoGenera/",
           filename = "genus_country_constrained_search")
```

```{r echo=FALSE, warning=FALSE}
library(taxize)
fam <- c("Aristolochiaceae", "Lecythidaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam, messages = FALSE)))

res <- powoGenera(powocodes$family, powocodes$uri,
                  genus = c("Asarum", "Bertholletia"),
                  country = c("Argentina", "Brazil", "French Guiana"),
                  verbose = TRUE,
                  save = FALSE)

knitr::kable(res[-c(7:10)])
```
