---
title: "Mining the top richest genera of flowering plants"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining the top richest genera of flowering plants}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.table{
  width: auto;
  font-size: 14px;
}
.table caption {
  font-size: 1em;
}
```

Here in this article, we show how to use the package's function `topGen` for 
mining the top most species rich genera for any family of flowering plants.\

\

# Setup

Install the latest development version of __expowo__ from 
[GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("DBOSlab/expowo")
```
```{r package load}
library(expowo)
```

\

# Mining the top richest genera for any plant family

The function `topGen` is relatively similar to the `megaGen`, but instead of 
using a specific threshold of species number to be considered a big genera, it 
saves a CSV file listing the top most species-rich genera of any target 
angiosperm family, and their associated number of accepted species based on an
integer number set in the argument limit. In the example below, we used the 
default to search for the top ten richest genera (i.e., when the argument limit
is set as NULL, the function searches for the 10 most diverse genera within a 
plant family). Some columns were omitted to display the results, and the rows 
show the top ten of the three chosen families: Bignoniaceae, Solanaceae and 
Lecythidaceae. Note that the table does not have 30 rows, what would be expected 
if we multiply each family (n = 3) by ten. It is not an error, since some plant 
families do not have many genera, which is the case of Begoniaceae with 2 
genera.\

```{r, eval = FALSE}
ABL_top <- topGen(family = c("Aristolochiaceae", "Begoniaceae", "Lecythidaceae"),
                  limit = NULL,
                  verbose = TRUE,
                  save = FALSE,
                  dir = "results_topGen",
                  filename = "Aristo_Bego_Lecythidaceae_search")
```

```{r, echo = FALSE, warning = FALSE}
library(dplyr, warn.conflicts = FALSE)
utils::data("angioGenera")
family <- c("Begoniaceae", "Aristolochiaceae", "Lecythidaceae")

genus <- angioGenera$genus[angioGenera$family %in% family]
species_number <- angioGenera$species_number[angioGenera$family %in% family]
family <- angioGenera$family[angioGenera$family %in% family]
table <- data.frame(family, genus, species_number)
res <- table %>% arrange(desc(table$species_number)) %>% group_by(family) %>% 
  slice(1:10)

knitr::kable(res,
             row.names = FALSE,
             align = 'c',
             caption = "TABLE 1. A general `topGen` search to mining the top 
             most species rich genera for some specific angiosperm 
             families.")
```

\

# Mining the top richest genera accross all angiosperm families

To mine a global checklist of the top species-richest genera for all 
families of flowering plants, including their associated species number, we 
recommend to load the dataframe-formatted data object called `POWOcodes` that 
comes associated with the __expowo__ package. The `POWOcodes` data 
object already contains the URI addresses for all angiosperms families 
recognized in the [POWO](https://powo.science.kew.org) database, so you just 
need to call it to your R environment.\

The example below shows how to mine all top most species-rich genera of
flowering plants by using the vector of all angiosperm families, the associated 
URI addresses stored in the `POWOcodes` object and the limit of 10 genera for 
each family.\

```{r, eval = FALSE}
data(POWOcodes)

ALL_top <- topGen(POWOcodes$family,
                  limit = 10,
                  verbose = TRUE,
                  save = FALSE,
                  dir = "results_topGen",
                  filename = "all_toprichest_angiosperm_genera")
```
