---
title: "Mining genus list for any plant family"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining genus list for any plant family}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The main goal of the __expowo__ package is to retrieve information about the 
diversity and distribution of any plant family as publicly available at the 
taxonomically verified database 
[Plants of the World Online (POWO)](https://powo.science.kew.org).\

The package is intended to efficiently mine the content within the source HTML 
pages for any specific genus and family. It can return a comma-separated values 
(CSV) file with the number of accepted species and country-level distribution 
for any genus as well as the full checklist of accepted species in any genus or 
family, their authorship, original publication and global distribution.\

Here in this article, we show how to use the package's function `powoGenera` for 
mining all accepted genera for any family of flowering plants.

# Setup

__expowo__ is not on CRAN yet but you can install the latest development version 
from [GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("DBOSlab/expowo")
```
```{r package load}
library(expowo)
```

# Mining all accepted genera for any angiosperm family

The function `powoGenera` returns a dataframe or saves a CSV file listing all 
genera with associated number of accepted species and their global geographic 
distribution at country or botanical level. The global classification of 
botanical divisions follows the 
[World Geographical Scheme](https://www.tdwg.org/standards/wgsrpd/)
for Recording Plant Distributions, which is already associated with each taxon's 
distribution in the POWO.

```{r, eval = FALSE}
library(taxize)
fam <- c("Aristolochiaceae", "Begoniaceae", "Martyniaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

powoGenera(powocodes$family, powocodes$uri,
           verbose = FALSE,
           save = TRUE,
           dir = "results_powoGenera/",
           filename = "Arist_Begon_Martyniaceae_search")
```

```{r, echo = FALSE, warning = FALSE}
data("POWOcodes")
fam <- c("Aristolochiaceae", "Begoniaceae", "Martyniaceae")
powocodes <- data.frame(family = fam,
                        uri = POWOcodes$uri[POWOcodes$family %in% fam])

res <- powoGenera(powocodes$family, powocodes$uri,
                  verbose = FALSE,
                  save = FALSE)

knitr::kable(res[-c(7:10)],
             caption = "TABLE 1. A general `powoGenera` search for mining all 
             accepted genera for some specific angiosperm families.")
```

# Mining all accepted genera for all angiosperm families

To mine a global genus checklist with associated species number and distribution 
for all families of flowering plants, we recommend to
load the dataframe-formatted data object called `POWOcodes` that comes 
associated with the __expowo__ package. Because the `POWOcodes` data object 
already contains the URI addresses for all angiosperms families recognized in 
the [POWO](https://powo.science.kew.org) database, you do not need to perform 
the time-consuming URI search using __taxize__. \

The example below shows how to mine a global checklist of all accepted genera of
flowering plants by using the vector of all angiosperm families and associated 
URI addresses stored in the `POWOcodes` object.

```{r, eval = FALSE}
data(POWOcodes)

powoGenera(POWOcodes$family, POWOcodes$uri,
           verbose = TRUE,
           save = TRUE,
           dir = "results_powoGenera/",
           filename = "all_angiosperm_genera")
```

# Narrowing down the `powoGenera` search based on a specified country vector

You can also narrow down the search for all accepted genera of any family so as 
to focus on just a particular country or a list of countries. You just need to 
define a vector of country names in the argument \code{country}. In the example 
below, see that we have originally searched for the genera within the families 
__Aristolochiaceae__, __Lecythidaceae__, __Fagaceae__, and __Dipterocarpaceae__, 
but the function only returned a list of genera of __Aristolochiaceae__ and 
__Lecythidaceae__, because the remaining families do not have any genera 
recorded in the list of countries of interest, i.e. __Argentina__, __Brazil__ 
or __French Guiana__.

```{r, eval = FALSE}
library(taxize)
fam <- c("Aristolochiaceae", "Dipterocarpaceae", "Fagaceae", "Lecythidaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

powoGenera(powocodes$family, powocodes$uri,
           country = c("Argentina", "Brazil", "French Guiana"),
           verbose = FALSE,
           save = TRUE,
           dir = "results_powoGenera/",
           filename = "country_constrained_search")
```

```{r, echo = FALSE, warning = FALSE}
fam <- c("Aristolochiaceae", "Dipterocarpaceae", "Fagaceae", "Lecythidaceae")
powocodes <- data.frame(family = fam,
                        uri = POWOcodes$uri[POWOcodes$family %in% fam])

res <- powoGenera(powocodes$family, powocodes$uri,
                  country = c("Argentina", "Brazil", "French Guiana"),
                  verbose = FALSE,
                  save = FALSE)

knitr::kable(res[-c(7:10)],
             caption = "TABLE 2. A `powoGenera` search based on a specified 
             country vector.")
```


# Narrowing down the `powoGenera` search based on a specified genus vector

You may want to retrieve information for just one or a list of accepted 
genera from a given country (or from a list of countries). Just like before,
you only need to define a vector of genus names in the argument \code{genus} and
a vector of country names in the argument \code{country}. In the example below,
see that we have again searched for just the genera ***Asarum*** and 
***Bertholletia*** of the families __Aristolochiaceae__ and __Lecythidaceae__, 
but the function only returned the Lecythidaceae genus ***Bertholletia***, 
because ***Asarum*** does not occur in any of the provided list of countries, 
i.e. __Argentina__, __Brazil__ or __French Guiana__.

```{r, eval = FALSE}
library(taxize)
fam <- c("Aristolochiaceae", "Lecythidaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

powoGenera(powocodes$family, powocodes$uri,
           genus = c("Asarum", "Bertholletia"),
           country = c("Argentina", "Brazil", "French Guiana"),
           verbose = TRUE,
           save = TRUE,
           dir = "results_powoGenera/",
           filename = "genus_country_constrained_search")
```

```{r echo=FALSE, warning=FALSE}
res <- powoGenera(powocodes$family[c(1,4)], powocodes$uri[c(1,4)],
                  genus = c("Asarum", "Bertholletia"),
                  country = c("Argentina", "Brazil", "French Guiana"),
                  verbose = TRUE,
                  save = FALSE)

knitr::kable(res[-c(7:10)], 
             caption = "TABLE 3. A `powoGenera` search based on specified genus 
             and country vectors.")
```
