---
title: "Mining species number of any plant family"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining species number of any plant family}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The main goal of the __expowo__ package is to retrieve information about the 
diversity and distribution of any plant family as publicly available at the 
taxonomically verified database 
[Plants of the World Online (POWO)](https://powo.science.kew.org).\

The package is intended to efficiently mine the content within the source HTML 
pages for any specific genus and family. It can return a comma-separated values 
(CSV) file with the number of accepted species and country-level distribution 
for any genus as well as the full checklist of accepted species in any genus or 
family, their authorship, original publication and global distribution.\

Here in this article, we show how to use the package's function `powoFam` 
for mining the species number for any family of flowering plants. To a provided
family name or list of names, the `powoFam` visit each one of the genus pages 
at POWO' website, extracts the accepted species number and then sum these values
to deliver the total of species within each plant family.\


# Setup

__expowo__ is not on CRAN yet but you can install the latest development version 
from [GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("DBOSlab/expowo")
```
```{r package load}
library(expowo)
```

# Mining the total species number for any angiosperm family

The function `powoFam` returns a dataframe or saves a CSV file listing the total
species number for any family, excluding hybrid species from the count, their genus, publication and authorship. The global classification of botanical divisions follows the 
[World Geographical Scheme](https://www.tdwg.org/standards/wgsrpd/) for 
Recording Plant Distributions, which is already associated with each taxon's 
distribution in the POWO.\

The example below shows how to mine the distribution for a specified vector of 
two plant families, Araceae and Sapotaceae.\
 
```{r, eval = FALSE}
library(taxize)
fam <- c("Araceae", "Sapotaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

powoFam(powocodes$family, powocodes$uri,
        verbose = TRUE,
        save = TRUE,
        dir = "results_powoFam/",
        filename = "Araceae_Sapotaceae")
```

```{r, echo = FALSE, warning = FALSE}
data("POWOcodes")
fam <- c("Araceae", "Sapotaceae")
powocodes <- data.frame(family = fam,
                        uri = POWOcodes$uri[POWOcodes$family %in% fam])

res <- powoFam(powocodes$family, powocodes$uri,
               verbose = FALSE,
               save = FALSE,
               dir = "results_powoFam/",
               filename = "Araceae_Sapotaceae")
knitr::kable(res,
             align = 'c',
             caption = "TABLE 1. A general `powoFam` search for mining 
             species number of the Araceae and Sapotaceae families.")
```
# Mining species number accross all angiosperm families

To mine the species number of all families of flowering 
plants, we recommend to load the 
data frame-formatted object called `POWOcodes` that comes associated with
the __expowo__ package. Because the `POWOcodes` data object already contains the
URI addresses for all angiosperms families recognized in the 
[POWO](https://powo.science.kew.org) database, you do not need to perform the 
time-consuming URI search using __taxize__. \

The example below shows how to mine the species number of all plant families by using the vector of all angiosperm families and associated URI addresses stored in the `POWOcodes` object.

```{r, eval = FALSE}
data(POWOcodes)

powoFam(POWOcodes$family, POWOcodes$uri,
        verbose = TRUE,
        save = TRUE,
        dir = "results_powoFam/",
        filename = "all_angiosperms_species_number")
```
