---
title: "Mining species list for any plant family"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining species list for any plant family}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.table{
  width: auto;
  font-size: 14px;
}
.table caption {
  font-size: 1em;
}
```

The main goal of the __expowo__ package is to retrieve information about the 
diversity and distribution of any plant family as publicly available at the 
taxonomically verified database 
[Plants of the World Online (POWO)](https://powo.science.kew.org).\

The package is intended to efficiently mine the content within the source HTML 
pages for any specific genus and family. It can return a comma-separated values 
(CSV) file with the number of accepted species and country-level distribution 
for any genus as well as the full checklist of accepted species in any genus or 
family, their authorship, original publication and global distribution.\

Here in this article, we show how to use the package's function `powoSpecies` 
for mining all accepted species for any genus or family of flowering plants.\

\

# Setup

__expowo__ is not on CRAN yet but you can install the latest development version 
from [GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("DBOSlab/expowo")
```
```{r package load}
library(expowo)
```

\

# Mining all accepted species for any angiosperm family

The function `powoSpecies` returns a dataframe or saves a CSV file listing all 
accepted species (including hybrid species or not), their publication, 
authorship, and global geographic distribution at country or botanical level for
any family of flowering plants. The global classification of botanical divisions
follows the [World Geographical Scheme](https://www.tdwg.org/standards/wgsrpd/) 
for Recording Plant Distributions, which is already associated with each taxon's
distribution in the POWO.\

The example below shows how to mine all accepted species for a specified vector 
of families. The output shown here (TABLE 1) is a simplified version, where we have removed some columns (genus, epithet and authors) so as to focus on the display of the taxonomic and distribution data.\

```{r, eval = FALSE}
library(taxize)

fam <- c("Amborellaceae", "Cabombaceae", "Martyniaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

res <- powoSpecies(powocodes$family, powocodes$uri,
                   hybridspp = FALSE,
                   verbose = TRUE,
                   save = TRUE,
                   dir = "results_powoSpecies/",
                   filename = "Ambor_Cabom_Martyniaceae_search")
```

```{r, echo = FALSE, warning = FALSE}
fam <- c("Amborellaceae", "Cabombaceae", "Martyniaceae")
utils::data("angioData")
df <- angioData[angioData$family %in% fam, ]

knitr::kable(df[-c(2, 3, 4, 5, 10, 11, 13)],
             row.names = FALSE,
             caption = "TABLE 1. A general `powoSpecies` search for mining all 
             accepted species for some specific angiosperm families.")
```

\

# Narrowing down the `powoSpecies` search based on a specified country vector

You can also narrow down the species checklist search of any family by
focusing on just a particular country or a list of countries. You just need to 
define a vector of country names in the argument \code{country}. In the example 
below, note that we have originally searched for the species within the families 
__Amborellaceae__, __Cabombaceae__, and __Martyniaceae__, but the function 
returned a smaller species list, because many species in the searched families 
are not recorded in any of the specified vector of country, i.e. __Brazil__ or 
__Colombia__.\

```{r, eval = FALSE}
library(taxize)

fam <- c("Amborellaceae", "Cabombaceae", "Martyniaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

powoSpecies(powocodes$family, powocodes$uri,
            hybridspp = FALSE,
            country = c("Brazil", "Colombia"),
            verbose = TRUE,
            save = F,
            dir = "results_powoSpecies/",
            filename = "country_constrained_search")
```

```{r, echo = FALSE, warning = FALSE}
utils::data("angioData")
fam <- c("Amborellaceae", "Cabombaceae", "Martyniaceae")
df <- angioData[angioData$family %in% fam, ]
country <- c("Brazil", "Colombia")
df <- df[df$native_to_country %in% country, ]

knitr::kable(df[-c(2, 3, 4, 5, 10, 11, 12, 13)],
             row.names = FALSE,
             caption = "TABLE 2. A `powoSpecies` search based on a specified 
             country vector.")
```

\

# Narrowing down the `powoSpecies` search based on a specified genus and country vector

You may want to retrieve information for just one or a list of accepted 
genera from a given country (or from a list of countries). Just like before,
you only need to define a vector of genus names in the argument \code{genus} and
a vector of country names in the argument \code{country}. In the example below,
see that we have searched for just the genera ***Amborella*** and 
***Cabomba*** of the families __Amborellaceae__ and __Cabombaceae__, 
but the function only returned the __Cabombaceae__ genus ***Cabomba***, 
because ***Amborella*** does not occur in any of the provided list of countries,
i.e. __Argentina__, __French Guiana__, or __Brazil__.\

```{r, eval = FALSE}
library(taxize)

fam <- c("Amborellaceae", "Cabombaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

powoSpecies(powocodes$family, powocodes$uri,
            genus = c("Amborella", "Cabomba"),
            hybridspp = FALSE,
            country = c("Argentina", "French Guiana", "Brazil"),
            verbose = TRUE,
            save = F,
            dir = "results_powoSpecies/",
            filename = "country_and_genus_constrained_search")
```

```{r, echo = FALSE, warning = FALSE}
utils::data("angioData")
genus <- c("Amborella", "Cabomba")
country <- c("Argentina", "French Guiana", "Brazil")
df <- angioData[angioData$genus %in% genus, ]
df <- df[df$native_to_country %in% country, ]

knitr::kable(df[-c(2, 3, 4, 5, 10, 11, 13)],
             row.names = FALSE,
             caption = "TABLE 3. A `powoSpecies` search based on specified genus
             and country vectors.")
```
\

# Mining all accepted species for all angiosperm families

To mine a species checklist for all families of flowering plants, you recommend 
to load the dataframe-formatted data object called `POWOcodes` that comes 
associated with the __expowo__ package. Because the `POWOcodes` data object 
already contains the URI addresses for all angiosperms families recognized in 
the [POWO](https://powo.science.kew.org) database, you do not need to perform 
the time-consuming URI search using __taxize__. \

The example below shows how to mine a global checklist of all accepted species 
of flowering plants by using the vector of all angiosperm families and 
associated URI addresses stored in the `POWOcodes` object.\

```{r, eval = FALSE}
utils::data("POWOcodes")

powoSpecies(POWOcodes$family, POWOcodes$uri,
            hybridspp = TRUE,
            verbose = TRUE,
            save = TRUE,
            dir = "results_powoSpecies/",
            filename = "all_angiosperm_species")
```
