---
title: "Mining the top ten richest genera of flowering plants"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining the top ten richest genera of flowering plants}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The main goal of the __expowo__ package is to retrieve information about the 
diversity and distribution of any plant family as publicly available at the 
taxonomically verified database [Plants of the World Online (POWO)](https://powo.science.kew.org).\

The package is intended to efficiently mine the content within the source html pages
for any specific genus and family. It can return a comma-separated values (CSV) 
file with the number of accepted species and country-level distribution for any genus
as well as the full checklist of accepted species in any genus or family, their 
authorship, original publication and global distribution.\

Here in this article, we show how to use the package's function `toptenGen` for 
mining the top ten most species rich genera for any family of flowering plants. 

# Setup

__expowo__ is not on CRAN yet but you can install the latest development version 
from [GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("deborazuanny/expowo")
```
```{r package load}
library(expowo)
```

# Mining the top ten richest genera for any angiosperm family

The function `toptenGen` is relatively similar to the `megaGen`, but instead of 
using any specific threshold, it saves a CSV file listing the top ten most biodiverse
genera of any target angiosperm family, and their associated number of accepted species.

```{r, eval = FALSE}
library(taxize)
fam <- c("Bignoniaceae", "Lecythidaceae", "Sapotaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

toptenGen(powocodes$family, powocodes$uri,
          verbose = TRUE,
          save = TRUE,
          dir = "results_toptenGen/",
          filename = "Bigno_Lecyt_Sapotaceae_search")
```

```{r, echo = FALSE, warning = FALSE}
library(taxize)
fam <- c("Bignoniaceae", "Lecythidaceae", "Sapotaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam, messages = FALSE)))

res <- toptenGen(powocodes$family, powocodes$uri,
                 verbose = FALSE,
                 save = FALSE)

knitr::kable(res,
             caption = "TABLE 1. A general `toptenGen` search to mining the top 
             ten most species rich genera for some specific angiosperm families.")
```

# Mining the top ten richest genera accross all angiosperm families

To mine a global checklist of the top ten species-richest genera for all families 
of flowering plants, including their associated species number, we recommend to 
load the dataframe-formatted data object called `POWOcodes` that comes associated 
with the __expowo__ package. Because the `POWOcodes` data object already contains the 
URI addresses for all angiosperms families recognized in the [POWO](https://powo.science.kew.org) 
database, you do not need to perform the time-consuming URI search using __taxize__. \

The example below shows how to mine all top ten most species-rich genera of
flowering plants by using the vector of all angiosperm families and associated 
URI addresses stored in the `POWOcodes` object.

```{r, eval = FALSE}
data(POWOcodes)

toptenGen(POWOcodes$family, POWOcodes$uri,
          verbose = TRUE,
          save = TRUE,
          dir = "results_toptenGen/",
          filename = "all_toptenrichest_angiosperm_genera")
```
