---
title: "Mining species number of any plant family"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining species number of any plant family}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The main goal of the __expowo__ package is to retrieve information about the 
diversity and distribution of any plant family as publicly available at the 
taxonomically verified database 
[Plants of the World Online (POWO)](https://powo.science.kew.org).\

The package is intended to efficiently mine the content within the source HTML 
pages for any specific genus and family. It can return a comma-separated values 
(CSV) file with the number of accepted species and country-level distribution 
for any genus as well as the full checklist of accepted species in any genus or 
family, their authorship, original publication and global distribution.\

Here in this article, we show how to use the package's function `powoFam` 
for mining the species number for any family of flowering plants. To a provided
family name or list of names, the `powoFam` visit each one of the genus pages 
at POWO' website, extracts the accepted species number and then sum these values
to deliver the total of species within each plant family.\


# Setup

__expowo__ is not on CRAN yet but you can install the latest development version 
from [GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("DBOSlab/expowo")
```
```{r package load}
library(expowo)
```

# Mining the total species number for any angiosperm family

The function `powoFam` returns a dataframe or saves a CSV file listing the total
species number for any of the given family name(s),  (excluding hybrid species), their genus, publication 
and authorship. The global classification of botanical divisions follows the 
[World Geographical Scheme](https://www.tdwg.org/standards/wgsrpd/) for 
Recording Plant Distributions, which is already associated with each taxon's 
distribution in the POWO.\

The example below shows how to mine the distribution for a specified vector of 
two species. The output shown here (a table) is simplified by removing the 
family and distribution according to botanical subdivision columns.

```{r, eval = FALSE}
library(taxize)
fam <- c("Araceae", "Lecythidaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

res <- (powocodes$family, powocodes$uri,
        species = c("Anthurium longicaudatum", "Lecythis pisonis"),
        verbose = TRUE,
        save = TRUE,
        dir = "results_powoSpDist/",
        filename = "Araceae_Lecythidaceae")
```

```{r, echo = FALSE, warning = FALSE}
data("POWOcodes")
fam <- c("Araceae", "Lecythidaceae")
powocodes <- data.frame(family = fam,
                        uri = POWOcodes$uri[POWOcodes$family %in% fam])

res <- powoSpDist(powocodes$family, powocodes$uri,
                  species = c("Anthurium longicaudatum", "Lecythis pisonis"),
                  verbose = FALSE,
                  save = FALSE)

knitr::kable(res[-c(1, 3, 6, 9, 11, 13)],
             align = 'c',
             caption = "TABLE 1. A general `powoSpDist` search for mining 
             distribution for two angiosperm species.")
```
