---
title: "Mining distribution for any plant species"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining distribution for any plant species}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The main goal of the __expowo__ package is to retrieve information about the 
diversity and distribution of any plant family as publicly available at the 
taxonomically verified database 
[Plants of the World Online (POWO)](https://powo.science.kew.org).\

The package is intended to efficiently mine the content within the source HTML 
pages for any specific genus and family. It can return a comma-separated values 
(CSV) file with the number of accepted species and country-level distribution 
for any genus as well as the full checklist of accepted species in any genus or 
family, their authorship, original publication and global distribution.\

Here in this article, we show how to use the package's function `powoSpDist` 
for mining the entire distribution for any species of flowering plants. The 
`powoSpDist` uses the argument `species` to set the name or list of species 
names, and then the function access the auxiliary functions `getGenURI` and 
`getDist` to catch their respective distribution. 


# Setup

__expowo__ is not on CRAN yet but you can install the latest development version 
from [GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("DBOSlab/expowo")
```
```{r package load}
library(expowo)
```

# Mining the distribution for any angiosperm species

The function `powoSpDist` returns a dataframe or saves a CSV file listing the 
entire global distribution at country or botanical level for any of the given 
species scientific name(s) (excluding hybrid species), their genus, publication 
and authorship. The global classification of botanical divisions follows the 
[World Geographical Scheme](https://www.tdwg.org/standards/wgsrpd/) for 
Recording Plant Distributions, which is already associated with each taxon's 
distribution in the POWO.\

The example below shows how to mine the distribution for a specified vector of 
two species. The output shown here (a table) is simplified by removing the 
family and distribution according to botanical subdivision columns.

```{r, eval = FALSE}
library(taxize)
fam <- c("Araceae", "Lecythidaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

res <- (powocodes$family, powocodes$uri,
        species = c("Anthurium longicaudatum", "Lecythis pisonis"),
        verbose = TRUE,
        save = TRUE,
        dir = "results_powoSpDist/",
        filename = "Araceae_Lecythidaceae")
```

```{r, echo = FALSE, warning = FALSE}
data("POWOcodes")
fam <- c("Araceae", "Lecythidaceae")
powocodes <- data.frame(family = fam,
                        uri = POWOcodes$uri[POWOcodes$family %in% fam])

res <- powoSpDist(powocodes$family, powocodes$uri,
                  species = c("Anthurium longicaudatum", "Lecythis pisonis"),
                  verbose = FALSE,
                  save = FALSE)

knitr::kable(res[-c(1, 3, 6, 9, 11, 13)],
             align = 'c',
             caption = "TABLE 1. A general `powoSpDist` search for mining 
             distribution for two angiosperm species.")
```

# Mining distribution for all angiosperm species

To mine the distribution checklist for all accepted species of flowering plants, 
you can load the dataframe-formatted data object called `POWOcodes` 
that comes associated with the __expowo__ package. Because the `POWOcodes` data 
object already contains the URI addresses for all angiosperms families 
recognized in the [POWO](https://powo.science.kew.org) database, you do not need 
to perform the time-consuming URI search using __taxize__. Otherwise, we 
strictly suggest you to perform this task in parallel. The parallelization 
strategy used in the further example is multissesion (recommended if you are
using a computer with Windows OS).\

The example below shows how to mine the global distribution of all accepted non
hybrid species of flowering plants by using the vector of all angiosperm 
families and associated URI addresses stored in the `POWOcodes` object.

```{r, eval = FALSE}
library(future)
library(future.apply)

plan(multisession)
data(POWOcodes)

resSpDist %<-% powoSpDist(POWOcodes$family, POWOcodes$uri,
                          verbose = TRUE,
                          save = TRUE,
                          dir = "results_powoSpDist/",
                          filename = "all_angiosperm_distribution")
resSpDist
plan(sequential)
```
