---
title: "Mining megadiverse genera of flowering plants"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining megadiverse genera of flowering plants}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.table{
  width: auto;
  font-size: 14px;
}
.table caption {
  font-size: 1em;
}
```

The main goal of the __expowo__ package is to retrieve information about the 
diversity and distribution of any plant family as publicly available at the 
taxonomically verified database 
[Plants of the World Online (POWO)](https://powo.science.kew.org).\

The package is intended to efficiently mine the content within the source HTML 
pages for any specific genus and family. It can return a comma-separated values 
(CSV) file with the number of accepted species and country-level distribution 
for any genus as well as the full checklist of accepted species in any genus or 
family, their authorship, original publication and global distribution.\

Here in this article, we show how to use the package's function `megaGen` for 
mining the megadiverse genera for any family of flowering plants. 
[Frodin (2004)](https://onlinelibrary.wiley.com/doi/abs/10.2307/4135449) suggests 500 species as the threshold for a genus to be considered megadiverse. The function `megaGen` 
uses any integer numeric value in the argument `thld` to define the threshold 
for a megadiverse genus.\

\

# Setup

__expowo__ is not on CRAN yet but you can install the latest development version 
from [GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("DBOSlab/expowo")
```
```{r package load}
library(expowo)
```

\

# Mining megadiverse genera for any angiosperm family

The function `megaGen` uses a specified threshold number to return a dataframe 
or save a CSV file listing all megadiverse genera for any family of flowering 
plants and their associated number of accepted species. In this example, we 
performed a search for three plant families using a threshold of 500 species, 
but the output only contains results for **Aristolochia** and **Begonia** of __Aristolochiaceae__ and __Begoniaceae__ families.\

```{r, eval = FALSE}
library(taxize)

fam <- c("Aristolochiaceae", "Begoniaceae", "Martyniaceae")
powocodes <- cbind(family = fam,
                   data.frame(taxize::get_pow(fam)))

megaGen(powocodes$family, powocodes$uri,
        thld = 500,
        verbose = TRUE,
        save = TRUE,
        dir = "results_megaGen/",
        filename = "Arist_Begon_Piper_Solanaceae_search")
```

```{r, echo = FALSE, warning = FALSE}
utils::data("angioGenera")
family <- c("Aristolochiaceae", "Begoniaceae")
genus <- c("Aristolochia", "Begonia")
species_number <- angioGenera$species_number[angioGenera$genus %in% genus]
authors <- c("L.", "L.")
scientific_name <- c("Aristolochia L.", "Begonia L.")
powo_uri <- angioGenera$powo_uri[angioGenera$genus %in% genus]
kew_id <- angioGenera$kew_id[angioGenera$genus %in% genus]

res <- data.frame(family, genus, authors, scientific_name,
                  species_number, kew_id, powo_uri)

knitr::kable(res,
             row.names = FALSE,
             caption = "TABLE 1. A general `megaGen` search to mining the
             megadiverse genera for some specific angiosperm families,
             based on a specified threshold of 500 species.")
```

\

# Mining megadiverse genera accross all angiosperm families

To mine a global checklist of megadiverse genera for all families of flowering 
plants, including their associated species number, we recommend to load the 
dataframe-formatted data object called `POWOcodes` that comes associated with
the __expowo__ package. Because the `POWOcodes` data object already contains the 
URI addresses for all angiosperms families recognized in the 
[POWO](https://powo.science.kew.org) database, you do not need to perform the 
time-consuming URI search using __taxize__.\

The example below shows how to mine a global checklist of all most speciose 
genera of flowering plants by using the vector of all angiosperm families and 
associated URI addresses stored in the `POWOcodes` object.\

```{r, eval = FALSE}
data(POWOcodes)

megaGen(POWOcodes$family, POWOcodes$uri,
        thld = 500,
        verbose = TRUE,
        save = TRUE,
        dir = "results_megaGen/",
        filename = "all_megadiverse_angiosperm_genera")
```
