---
title: "Mining species number of any plant family"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining species number of any plant family}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.table{
  width: auto;
  font-size: 14px;
}
.table caption {
  font-size: 1em;
}
```

Here in this article, we show how to use the package's function `powoFam` 
for mining the species number for any family of flowering plants. By providing
a family name or a vector of family names, the `powoFam` function visits each 
POWO page of the genera within the queried family(ies) to extract the accepted 
species number. Then, it sums these values to deliver the total species number
within each plant family.\

\

# Setup

Install the latest development version of __expowo__ from 
[GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("DBOSlab/expowo")
```
```{r package load}
library(expowo)
```

\

# Mining the total species number for any angiosperm family

The function `powoFam` returns a dataframe or saves a CSV file listing the total
species number (excluding hybrid species) for the queried plant families.\

The example below shows how to mine the total species number by defining a vector of three families: __Cabombaceae__, __Lecythidaceae__, and __Martyniaceae__. The resulting output (TABLE 1) is also saved directly in the specified directory __results_powoFam__.\

```{r, eval = FALSE}
powoFam(family = c("Cabombaceae", "Lecythidaceae", "Martyniaceae"),
        verbose = TRUE,
        save = TRUE,
        dir = "results_powoFam/",
        filename = "Camb_Lecy_Martyniaceae_diversity")
```

```{r, echo = FALSE, warning = FALSE}
utils::data("angioData")
utils::data("POWOcodes")
family <- c("Cabombaceae", "Lecythidaceae", "Martyniaceae")
fams <- angioData$family[angioData$family %in% family]
sp_nbr <- as.data.frame(table(fams))
powo_uri <- gsub("^http", "https", POWOcodes$uri[POWOcodes$family %in% family])
kew_id <- gsub(".+[:]", "", powo_uri)
species_number <- sp_nbr$Freq
res <- data.frame(family, species_number, kew_id, powo_uri)

knitr::kable(res,
             row.names = FALSE,
             align = 'c',
             caption = "TABLE 1. A general `powoFam` search for mining the total species number of three angiosperm families.")
```

\

# Mining species number accross all angiosperm families

To mine the species number of all families of flowering plants, we recommend to 
load the data frame-formatted object called `POWOcodes` that comes associated with the __expowo__ package. The `POWOcodes` data object already contains the URI addresses for all angiosperms families recognized in the [POWO](https://powo.science.kew.org) database, so you just need to call it
to your R environment. \

The example below shows how to mine the species number of all plant families by 
using the vector of all angiosperm families and associated URI addresses stored 
in the `POWOcodes` object.\

```{r, eval = FALSE}
utils::data(POWOcodes)

powoFam(POWOcodes$family,
        verbose = TRUE,
        save = TRUE,
        dir = "results_powoFam/",
        filename = "all_angiosperms_species_number")
```
