---
title: "Mining distribution for any plant species"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining distribution for any plant species}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.table{
  width: auto;
  font-size: 14px;
}
.table caption {
  font-size: 1em;
}
```

Here in this article, we show how to use the package's function `powoSpDist` 
for mining the entire distribution for any species of flowering plants. The 
`powoSpDist` uses the argument `species` to set the name or list of species 
names, and then the function access the auxiliary functions `.getgenURI` and 
`getInfo` to extract their respective distribution.\

\

## Setup

Install the latest development version of __expowo__ from 
[GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("DBOSlab/expowo")
```
```{r package load, message=FALSE, warning=FALSE}
library(expowo)
```

\

## Mining the distribution for any plant species

The function `powoSpDist` returns a dataframe or saves a CSV file listing the 
entire global distribution at country or botanical level for any of the given 
species scientific name(s) (excluding hybrid species), their genus, publication 
and authorship. The global classification of botanical divisions follows the 
[World Geographical Scheme](https://www.tdwg.org/standards/wgsrpd/) for 
Recording Plant Distributions, which is already associated with each taxon's 
distribution at POWO.\

The example below shows how to mine the distribution for a specified vector of 
two species. The output shown here (a table) is simplified by removing the 
family and distribution according to botanical subdivision columns.\

```{r, eval = FALSE}
BL_dist <- powoSpDist(family = c("Begoniaceae", "Lecythidaceae"),
                      species = c("Hillebrandia sandwicensis", "Lecythis pisonis"),
                      verbose = TRUE,
                      save = FALSE,
                      dir = "results_powoSpDist",
                      filename = "Begoniaceae_Lecythidaceae")
```

\

```{r, echo = FALSE, warning = FALSE}
utils::data("angioData")
fam <- c("Begoniaceae", "Lecythidaceae")
df <- angioData[angioData$family %in% fam, ]
species <- c("Hillebrandia sandwicensis", "Lecythis pisonis")
df <- df[df$taxon_name %in% species, ]

knitr::kable(df[-c(2, 3, 6, 9, 11, 13)],
             row.names = FALSE,
             align = 'c',
             caption = "TABLE 1. A general `powoSpDist` search for mining 
             distribution of two angiosperm species.")
```

\

## Mining distribution for all angiosperm species

To mine the distribution checklist for all accepted species of flowering plants, 
you can load the dataframe-formatted data object called `POWOcodes` 
that comes associated with the __expowo__ package. The `POWOcodes` data 
object already contains the URI addresses for all angiosperms families 
recognized in the [POWO](https://powo.science.kew.org) database, you just need 
to call it to your R environment.\

The example below shows how to mine the global distribution of all accepted non
hybrid species of flowering plants by using the vector of all angiosperm 
families and associated URI addresses stored in the `POWOcodes` object. Note that
the argument `species` should be as `NULL`, so that the search is not constrained
to any particular species, i.e. the function will return global distribution for
all species in all queried families.\

```{r, eval = FALSE}
data(POWOcodes)

ALL_dist <- powoSpDist(POWOcodes$family,
                       species = NULL,
                       verbose = TRUE,
                       save = FALSE,
                       dir = "results_powoSpDist",
                       filename = "all_angiosperm_distribution")
```

\

## Reference

POWO (2019). "Plants of the World Online. Facilitated by the Royal Botanic Gardens, Kew. Published on the Internet; http://www.plantsoftheworldonline.org/ Retrieved April 2023."
