---
title: "Mining genus list for any plant family"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mining genus list for any plant family}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css, echo=FALSE}
.table{
  width: auto;
  font-size: 14px;
}
.table caption {
  font-size: 1em;
}
```

Here in this article, we show how to use the package's function `powoGenera` for
mining all accepted genera for any family of flowering plants.\

\

# Setup

Install the latest development version of __expowo__ from 
[GitHub](https://github.com/):

``` r
#install.packages("devtools")
devtools::install_github("DBOSlab/expowo")
```
```{r package load}
library(expowo)
```

\

# Mining all accepted genera for any angiosperm family

The function `powoGenera` returns a dataframe or saves a CSV file listing all 
genera with associated number of accepted species and their global geographic 
distribution at country or botanical level. The global classification of 
botanical divisions follows the 
[World Geographical Scheme](https://www.tdwg.org/standards/wgsrpd/)
for Recording Plant Distributions, which is already associated with each taxon's
distribution in POWO.\

```{r, eval = FALSE}
ABM <- powoGenera(family = c("Aristolochiaceae", "Begoniaceae", "Martyniaceae"),
                  verbose = FALSE,
                  save = FALSE,
                  dir = "results_powoGenera/",
                  filename = "Arist_Begon_Martyniaceae_search")
```

```{r, echo = FALSE, warning = FALSE}
utils::data("POWOcodes")

res <- powoGenera(family = c("Aristolochiaceae", "Begoniaceae", "Martyniaceae"),
                  verbose = FALSE,
                  save = FALSE)

knitr::kable(res[-c(7:10)],
             row.names = FALSE,
             caption = "TABLE 1. A general `powoGenera` search for mining all 
             accepted genera for some specific angiosperm families.")
```

\

# Mining all accepted genera for all angiosperm families

To mine a global genus checklist with associated species number and distribution 
for all families of flowering plants, we recommend to
load the dataframe-formatted data object called `POWOcodes` that comes 
associated with the __expowo__ package. Because the `POWOcodes` data object 
already contains the URI addresses for all angiosperms families recognized in 
the [POWO](https://powo.science.kew.org) database, you just need to run the code 
below.\

The example below shows how to mine a global checklist of all accepted genera of
flowering plants by using the vector of all angiosperm families and associated 
URI addresses stored in the `POWOcodes` object.\

```{r, eval = FALSE}
data(POWOcodes)

ALL_gen <- powoGenera(POWOcodes$family,
                      verbose = TRUE,
                      save = FALSE,
                      dir = "results_powoGenera/",
                      filename = "all_angiosperm_genera")
```

\

# Narrowing down the `powoGenera` search based on a specified country vector

You can also narrow down the search for all accepted genera of any family so as 
to focus on just a particular country or a list of countries. You just need to 
define a vector of country names in the argument `country`. In the example 
below, see that we have originally searched for the genera within the families Aristolochiaceae, Lecythidaceae, Fagaceae, and Dipterocarpaceae, 
but the function only returned a list of genera of Aristolochiaceae and 
Lecythidaceae, because the remaining families do not have any genera 
recorded in the list of countries of interest, i.e. Argentina, Brazil 
or French Guiana.\

```{r, eval = FALSE}
ADFL <- powoGenera(family = 
                     c("Aristolochiaceae", "Dipterocarpaceae", "Fagaceae", "Lecythidaceae"),
                   country = c("Argentina", "Brazil", "French Guiana"),
                   verbose = FALSE,
                   save = FALSE,
                   dir = "results_powoGenera/",
                   filename = "country_constrained_search")
  
```

```{r, echo = FALSE, warning = FALSE}
res <- powoGenera(family = c("Aristolochiaceae", "Dipterocarpaceae", "Fagaceae", "Lecythidaceae"),
                  country = c("Argentina", "Brazil", "French Guiana"),
                  verbose = FALSE,
                  save = FALSE)

knitr::kable(res[-c(7:12)],
             row.names = FALSE,
             caption = "TABLE 2. A `powoGenera` search based on a specified 
             country vector.")
```

\

# Narrowing down the `powoGenera` search based on a specified genus vector

You may want to retrieve information for just one or a list of accepted 
genera from a given country (or from a list of countries). Just like before,
you only need to define a vector of genus names in the argument `genus` and
a vector of country names in the argument `country`. In the example below,
see that we have again searched for just the genera ***Asarum*** and 
***Bertholletia*** of the families Aristolochiaceae and Lecythidaceae, 
but the function only returned the Lecythidaceae genus ***Bertholletia***, 
because ***Asarum*** does not occur in any of the provided list of countries, 
i.e. Brazil or French Guiana.\

```{r, eval = FALSE}
AL <- powoGenera(family = c("Aristolochiaceae", "Lecythidaceae"), 
                 genus = c("Asarum", "Bertholletia"),
                 country = c("Brazil", "French Guiana"),
                 verbose = TRUE,
                 save = FALSE,
                 dir = "results_powoGenera/",
                 filename = "genus_country_constrained_search")
```

```{r, echo = FALSE, warning = FALSE}
res <- powoGenera(family = c("Aristolochiaceae", "Lecythidaceae"),
                  genus = c("Asarum", "Bertholletia"),
                  country = c("Brazil", "French Guiana"),
                  verbose = FALSE,
                  save = FALSE)

knitr::kable(res[-c(8:12)],
             row.names = FALSE,
             caption = "TABLE 3. A `powoGenera` search based on specified genus 
             and country vectors.")
```
